---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
## QUESTION 2
Question 2 tests how the UPDRS score is affected by one’s vocal qualities (Shimmer, PPE and HNR) and sex.

```{r}
library(dplyr)
library(car)
library(fmsb)
library(ALSM)
library(onewaytests)
library(MASS)
library(readxl)
library(fmsb)
library(boot)
library(leaps)
library(caret)

```



```{r}
data<-read.csv('C:/Users/Ankita Mishra/Desktop/parkinsons_data.csv', header =TRUE, sep=",")

#modify data
data_group= data
data_group$test_time= as.integer(data_group$test_time)
data_group <- data_group %>%
  group_by('subject#', age, sex, test_time) %>%
  dplyr::summarise_all("median")%>%
  as.data.frame()

data= data_group

#define variables
data$y = data$total_UPDRS
data$x5 = data$sex
data$x6= data$Shimmer
data$x8= data$HNR
data$x9 = data$NHR

#correlation between the variables
df= data.frame(data$x5, data$x6, data$x8, data$x9)
pairs(df)
cor(df)
```

The full model includes all the variables and the interaction terms. Running a Shapiro test and Brown-Forsythe test on the data tells us the data is non-normal and doesn't have constant variance as interpreted by the p-value.
```{r}
#full model

fullmod = lm(y~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data)

res= resid(fullmod)
shapiro.test(res)

library(onewaytests)
data$gr = cut(data$y, 5)
data$residual = fullmod$residuals
bf.test(residual~gr, data)
```

Checking for influential points, row 25 seems to be the outlier hence it is removed from the data.

```{r}
#remove the influential point
data= data[-c(25), ]
fullmod = lm(y~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data)

res = resid(fullmod)
shapiro.test(res)

library(onewaytests)
data$gr = cut(data$y, 5)
data$residual = fullmod$residuals
bf.test(residual~gr, data)

```

Since our model is still non-normal and doesn't have constant variance, we transform Y. This gives us a lambda value= 1.060606

```{r}
#find lambda and transform Y
library(MASS)
bcmle= boxcox(fullmod, lambda= seq(-3, 3, by=0.1))
lambda= bcmle$x[which.max(bcmle$y)]
lambda

data$newy= (data$y)**lambda

#modified model after transformation
modY= lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data)

res= resid(modY)
shapiro.test(res)

data$gr= cut(data$newy, 5)
data$residual = modY$residuals
bf.test(residual~gr, data)

```
To reduce the non-normality and non-contant variance, we attempted the weighted least squares method
```{r}
# weighted least squares method
wts1=1/fitted(lm(abs(residuals(modY))~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x8+as.factor(x5)*x8, data))^2
mod1=lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, weight=wts1, data=data)

res = resid(mod1)
shapiro.test(res)
library(onewaytests)
data$gr = cut(data$newy, 5)
data$residual = mod1$residuals
bf.test(residual~gr, data)

plot(fitted(mod1), residuals(mod1))

wts2=1/fitted(lm(abs(residuals(mod1))~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x8+as.factor(x5)*x8, data))^2
mod2=lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, weight=wts2, data=data)

res = resid(mod2)
shapiro.test(res)
data$gr = cut(data$newy, 5)
data$residual = mod2$residuals
bf.test(residual~gr, data)

plot(fitted(mod2), residuals(mod2))

wts3=1/fitted(lm(abs(residuals(mod2))~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x8+as.factor(x5)*x8, data))^2
mod3=lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, weight=wts2, data=data)

res = resid(mod3)
shapiro.test(res)
data$gr = cut(data$newy, 5)
data$residual = mod3$residuals
bf.test(residual~gr, data)

plot(fitted(mod3), residuals(mod3))
```
The WLS made the model worse, thus we continue with the transformed Y model.


```{r}
plot(fitted(modY), residuals(modY))
abline(0,0)
res= resid(modY)
qqnorm(res)
qqline(res)
```


```{r}
#bootstrapping
library(boot)

boot.reg <- function(data, indices, maxit=1000) {
  data <- data[indices,]
  data.mod<-lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data=data)

  coef(data.mod)
}

data_model_reg <- boot(data=data, statistic = boot.reg, R=1000, maxit=100)
data_model_reg
summary(modY)
```

```{r}
#cross-validation
library(MASS)
library(leaps)
library(caret)

set.seed(123)

train.control=trainControl(method="cv", number=10)
step.model=train(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data, method="leapBackward", tuneGrid=data.frame(nvmax=5), trControl=train.control)
step.model$results
summary(modY)
```
```{r}
fullmod= lm(newy~x9+x6+x8+as.factor(x5)+as.factor(x5)*x9+as.factor(x5)*x6+as.factor(x5)*x8, data)
redmod= lm(newy~as.factor(x5), data)
anova(redmod, fullmod)

qf(0.95, 6, 945)
```
Since Fs= 16.404 > F(0.95, 6, 945) = 2.108158, we reject H0 and hence vocal qualities like Shimmer, HNR, NHR and their interaction terms with sex have a significant impact on UPDRS score